---
title: "MAT1856/APM466 Assignment 1 Appendix for R code"
output: html_notebook
---

1. Import the data set & Select the 10 bonds out.
```{r}
library(readxl)
all_bonds = read_excel("results_modified.xlsx")
# View(results_modified)

selected_ISIN = paste("CA135087",
                      c("D929","E596","F254","F585","G328","ZU15","H490","A610","J546","J967", "K528"), sep="")
selected_bonds = all_bonds[all_bonds$ISIN %in% selected_ISIN,]
selected_bonds$ISIN = c("D929","E596","F254","F585","G328","ZU15","H490","A610","J546","J967", "K528")
# View(selected_bonds)

# attach(selected_bonds)

today = c(2,3,6,7,8,9,10,13,14,15)
# Modify id for ZU15 (6) and A610 (8).
modify_id = c(6, 8)

```


2. Calculate the DP for each bond i at each day t.
```{r}
t = 1
# Calculate each column DP_i for a given t.
while (t <= 10) {
  colname_DP = paste("DP", toString(t), sep="")
  
  # Calculate column DP_i at day t
  AI = (4*30 + (today[t]))/360 * selected_bonds[["Coupon"]] * 100
  DP = AI + selected_bonds[[paste("CP", toString(t), sep="")]]
  
  selected_bonds[[colname_DP]] = DP
  t = t + 1
}

# Modify for the ZU15 (6) & A610 (8).

t = 1
while (t<=10) {
  
  # Calculate each DP for bond i at day t.
  for (i in modify_id) {
    AI_i = (30+today[t])/360 * selected_bonds[["Coupon"]][i] * 100
    DP_i = AI_i + selected_bonds[[paste("CP", toString(t), sep="")]][i]
    
    selected_bonds[[paste("DP", toString(t), sep="")]][i] = DP_i
  }
  
  t = t + 1
}
```


3. Calculate the YTM for each bond i at each day t.
```{r}
library(jrvFinance)

# Calculate each column YTM_i for a given day t.
t = 1
while(t<=10){
  colname_YTM = paste("YTM",toString(t),sep="")
  n = (selected_bonds[["Term(M)"]]-2)/6 + 1
  
  # Calculate each element YTM_i,t for bond i at day t.
  i = 1
  YTM_t = c()
  while (i<=11) {
    cash_flow = c(-selected_bonds[[paste("DP", toString(t), sep="")]][i])
    pmt_time = c(0)
    # Create the cash flow and time for bond i at day t.
    m = 1
    while (m<=n[i]-1) {
      cash_flow = append(cash_flow, selected_bonds[["Coupon"]][i]*100/2)
      pmt_time = append(pmt_time, (60-(today[t]-1))/360+1/2)
      m = m + 1
    }
    cash_flow = append(cash_flow, selected_bonds[["Coupon"]][i]*100/2 + 100)
    pmt_time = append(pmt_time, (60-(today[t]-1))/360+(n[i]-1)/2)
    
    YTM_it = irr(cf = cash_flow, cf.freq = 2, comp.freq = Inf, cf.t = pmt_time)
    YTM_t = append(YTM_t, YTM_it)
    i = i + 1
  }
  
  selected_bonds[[colname_YTM]] = YTM_t
  t = t + 1
}
```


4. Calculate the spot rate r for each bond i at each day t.
```{r}
# Calculate the column r_i at each day t.
t = 1
while(t <= 10){
  colname_r = paste("r",toString(t),sep="")
  r_1 = - log(selected_bonds[[paste("DP",toString(t),sep="")]][1] / (50*selected_bonds[["Coupon"]][1] + 100))/((60-(today[t]-1))/360)
  r = c(r_1)
  
  # Calculate the column r_i for bond 1 to 5 at each day t.
  i = 2
  while(i <= 5){
    term = 50*selected_bonds[["Coupon"]][i]*exp(-r_1 * ((60-(today[t]-1))/360) )
    j = 2
    while(j<=i-1){
      term = term + 50*selected_bonds[["Coupon"]][i]*exp(-((j-1)/2+ ((60-(today[t]-1))/360) )*r[j])
      j = j + 1
    }
    r_i = - log( (selected_bonds[[colname_DP]][i] - term) / (50*selected_bonds[["Coupon"]][i] + 100)) / ((i-1)/2 + ((60-(today[t]-1))/360)) 
    r = append(r, r_i)
    i = i + 1
  }
  
  new_r = c()
  l = 1
  while (l < 5) {
    new_r = append(new_r, (r[l]+r[l+1])/2 )
    l = l + 1
  }
  
  # Estimate the short rate for bond 6 at day t via interpoliation.
  term = 50*selected_bonds[["Coupon"]][6]*exp(-new_r[1] * ((30*5-(today[t]-1))/360) )
  j = 2
  while(j<=4){
      term = term + 50*selected_bonds[["Coupon"]][6]*exp(-((j-1)/2 + ((30*5-(today[t]-1))/360) )*new_r[j])
      j = j + 1
    }
  r_6 = -log( (selected_bonds[[colname_DP]][6] - term) / (50*selected_bonds[["Coupon"]][6] + 100)) / (5/2 - ((30+(today[t]-1))/360))
  new_r = append(new_r, r_6)
  r = append(r, 2*r_6-r[5])
  
  # Calculate the short rate for bond 7 at day t.
  term = 50*selected_bonds[["Coupon"]][7]*exp(-r_1 * ((60-(today[t]-1))/360) )
  j = 2
  while(j<=7-1){
    term = term + 50*selected_bonds[["Coupon"]][7]*exp(-((j-1)/2+ ((60-(today[t]-1))/360) )*r[j])
    j = j + 1
  }
  r_7 = - log( (selected_bonds[[colname_DP]][7] - term) / (50*selected_bonds[["Coupon"]][7] + 100)) / ((7-1)/2 + ((60-(today[t]-1))/360)) 
  r = append(r, r_i)
  
  new_r = append(new_r, (r[6]+r[7])/2)
  # Estimate the short rate for bond 8 at day t.
  term = 50*selected_bonds[["Coupon"]][8]*exp(-new_r[1] * ((30*5-(today[t]-1))/360) )
  j = 2
  while(j<=6){
      term = term + 50*selected_bonds[["Coupon"]][8]*exp(-((j-1)/2 + ((30*5-(today[t]-1))/360) )*new_r[j])
      j = j + 1
    }
  r_8 = -log( (selected_bonds[[colname_DP]][8] - term) / (50*selected_bonds[["Coupon"]][8] + 100)) / (7/2 - ((30+(today[t]-1))/360))
  new_r = append(new_r, r_8)
  r = append(r, 2*r_8-r[7])
  
  # Calculate the short rate for bond 9 and 10 at day t.
  i = 9
  while(i <= 11){
    term = 50*selected_bonds[["Coupon"]][i]*exp(-r_1 * ((60-(today[t]-1))/360) )
    j = 2
    while(j<=i-1){
      term = term + 50*selected_bonds[["Coupon"]][i]*exp(-((j-1)/2+ ((60-(today[t]-1))/360) )*r[j])
      j = j + 1
    }
    r_i = - log( (selected_bonds[[colname_DP]][i] - term) / (50*selected_bonds[["Coupon"]][i] + 100)) / ((i-1)/2 + ((60-(today[t]-1))/360)) 
    r = append(r, r_i)
    i = i + 1
  }
  
  colname_r = paste("r", toString(t), sep="")
  real_r = append(r[1:5], append(r_6, append(r_7, append(r_8, r[9:11]))))
  selected_bonds[[colname_r]] = real_r
  
  t = t + 1
}
```

5 Calculate the forward rate for each bond i at each day t.
```{r}
# Calculate the column f_i at each day t.
t = 1
term = selected_bonds[["Term(M)"]]/12
while (t <= 10) {
  f_i = c()
  r_year1_hat = (selected_bonds[[paste("r", toString(t), sep="")]][2] + selected_bonds[[paste("r", toString(t), sep="")]][3])/2
  
  # Calculate the element f at each day t.
  i = 1
  while (i <= 11) {
    f = ( selected_bonds[[paste("r", toString(t), sep="")]][i] * (term[i] - today[t]/360) - r_year1_hat * (1 - today[t]/360) ) / (term[i]-1)
    f_i = append(f_i, f)
    
    i = i + 1
  }
    
  colname_f = paste("f", toString(t), sep="")
  selected_bonds[[colname_f]] = f_i

  t= t + 1
}
```

6. Save the data set for the future usage.
```{r}
write.csv(selected_bonds, "selected_bonds.csv")
```

7. Read the data for the second question.
```{r}
library(readxl)
selected_bonds = read_excel("selected_bonds.csv")
# View(selected_bonds)
```

8. Time Series for yield.
```{r}
# Calculate the time series X_i for bond i. 
i = 1
X_ytm = c()
while (i <= 10) {
  yield_i_j_1 = as.numeric(selected_bonds[i,27:35])
  yield_i_j = as.numeric(selected_bonds[i,26:34])
  X_ytm_i = log(yield_i_j_1/yield_i_j)
  
  X_ytm = append(X_ytm, X_ytm_i)
  i = i + 2
}
X_ytm = matrix(X_ytm, ncol=5, byrow=FALSE)

# View(X_ytm)
M_ytm = cov(X_ytm)

eigen_space_ytm = eigen(M_ytm) 
eigen_space_ytm$values
eigen_space_ytm$vectors
```

9. Times Seriese for forward rate. 
```{r}
# Estimate the forward rate annualy.
t = 1
annualized_f = c()
while (t <= 10) {
  f_i = selected_bonds[[paste("f", toString(t), sep = "")]]
  annualized_f_i = c((f_i[1:4] + f_i[2:5])/2, f_i[6], (f_i[7]-f_i[6]) * 2/3+f_i[6], f_i[8], (f_i[9]-f_i[8]) * 2/3+f_i[8], (f_i[9:10]+f_i[10:11])/2)
  
  annualized_f = append(annualized_f, annualized_f_i[4:10])
  t = t + 1
}

annualized_f = matrix(annualized_f,ncol=10,byrow=FALSE)
# View(annualized_f)

# Calculate the time series X_i for forward rate.
i = 1
X_f = c()
while (i <= 8) {
  f_i_j_1 = as.numeric(annualized_f[i, 2:10])
  f_i = as.numeric(annualized_f[i, 1:9])
  X_f_i = log(f_i_j_1/f_i)
  
  X_f = append(X_f, X_f_i)
  i = i + 2
}
X_f = matrix(X_f, ncol=4, byrow=FALSE)

# View(X_f)
M_f = cov(X_f)

eigen_space_f = eigen(M_f) 
eigen_space_f$values
eigen_space_f$vectors
```
